PANCHANGE_descriptives

### 2020-11-11 PANCHANGE_exclusion_descriptives.Rmd
Added cor plot and scatterplots for little/much worse gad retro prepan

I interpreted the presence of an NA in the retrospective score as an indicator of saying "no change" in their baseline scores. Hence, if this was the case, I imputed their baseline score. However, this interpretation could be wrong.

Also, I'm not sure what to do if there is an NA for their baseline sum score (nothing to duplicate). This was the case for ocir for some reason: the retro scores were provided for some by not the corresponding baseline score?

@@

#Correlation between retropsective impute and prepan

```{r Correlation prep chunk}
# Select items for correlation matrix
prepan_retro_items <- c(
    "gad.sum_score_prepan",
    "gad.sum_score_retro", 
    "ocir.sum_score_prepan",
    "ocir.sum_score_retro",
    "phq.sum_score_prepan",
    "phq.sum_score_retro"
)

#Create a data set with only the COPING samples
prepan_retro_dat <- dat[str_detect(dat$Sample, "NBR|GLAD|EDGI"), ]

#Check that RAMP has been removed
prepan_retro_dat %>% 
  freq(Sample)

@@

# Save the p values from the correlation matrix in an object
prepan_retro.p.matrix <- as.matrix(prepan_retro.corr.mat$tests)
```

```{r Correlation matrix of prepan base sum scores}
corrplot(prepan_retro.corr.matrix, 
         method = "color", # objects to represent the correlations on plot
         type = "lower", # only use the lower triangle of the matrix
         diag = FALSE, # do not show the correlations on the diagonal
         addgrid.col = NA,
         addCoef.col = "black", # colour for the correlation coefficients in the plot
         tl.cex = 0.8,
         tl.col = "black",
         col=colorRampPalette(c("dodgerblue4","white","firebrick4"))(200), # colours for correlations
         # Combine with significance
         p.mat = prepan_retro.p.matrix, # Matrix with p values
         sig.level = 0.01, # Choose significant level
         insig = "blank" # Nonsignificant correlations have no colour
         )
```

>>>>

@@

#Correlation between whole sample retro impute and prepan
##Create function to use for all of the corplots

```{r function to create upper and lower CI matrices}

cor.mtest <- function(mat, conf.level = 0.95){
mat <- as.matrix(mat)
n <- ncol(mat)
p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
diag(p.mat) <- 0
diag(lowCI.mat) <- diag(uppCI.mat) <- 1
for(i in 1:(n-1)){
for(j in (i+1):n){
tmp <- cor.test(mat[,i], mat[,j], conf.level = conf.level)
p.mat[i,j] <- p.mat[j,i] <- tmp$p.value
lowCI.mat[i,j] <- lowCI.mat[j,i] <- tmp$conf.int[1]
uppCI.mat[i,j] <- uppCI.mat[j,i] <- tmp$conf.int[2]
}
}
return(list(p.mat, lowCI.mat, uppCI.mat))
}

```{r Whole sample Correlation prep chunk}
# Select items for correlation matrix
prepan_retro_items <- c(
    "gad.sum_score_prepan",
    "gad.sum_score_retro", 
    "phq.sum_score_prepan",
    "phq.sum_score_retro"
)

#Create a data set with only the GLAD and EDGI
prepan_retro_dat <- dat[str_detect(dat$Sample, "GLAD|EDGI"), ]

#Check that RAMP and NBR has been removed
prepan_retro_dat %>% 
  freq(Sample)

@@

# Save the p values from the correlation matrix in an object
prepan_retro.p.matrix <- as.matrix(prepan_retro.corr.mat$tests)

res1 <- cor.ci(prepan_retro.corr.matrix)
lowci <- cor.plot.upperLowerCi(res1)
lowci[upper.tri(lowci)] = t(lowci)[upper.tri(lowci)]

upci <- cor.plot.upperLowerCi(res1)
upci[lower.tri(upci)] = t(upci)[lower.tri(lowci)]

conf <- paste0("[", format(lowci, digits=2), " - ", format(upci, digits=2), "]")

xs <- row(lowci)
ys <- (ncol(lowci)+1) - col(lowci)

xs[lower.tri(xs,diag=TRUE)] <- 0
ys[lower.tri(ys,diag=TRUE)] <- 0

```{r Whole sample correlation matrix of prepan retro sum scores}
corrplot(prepan_retro.corr.matrix, 
         method = "color", # objects to represent the correlations on plot
         type = "lower", # only use the lower triangle of the matrix
         diag = FALSE, # do not show the correlations on the diagonal
         addgrid.col = NA,
         addCoef.col = "black", # colour for the correlation coefficients in the plot
         tl.cex = 0.8,
         tl.col = "black",
         col=colorRampPalette(c("dodgerblue4","white","firebrick4"))(200), # colours for correlations
         # Combine with significance
         p.mat = prepan_retro.p.matrix, # Matrix with p values
         sig.level = 0.01, # Choose significant level
         insig = "blank" # Nonsignificant correlations have no colour
         )

text(xs, ys, conf, pos=1, cex=0.8)

```{r Whole sample scatterplot for gad prepan base}

#Remove those saying "no different" since pandemic

whole_retro_prepan_scatter <- dat %>% 
  filter(gad.pandemic_felt_feelings == "Much worse" |
           gad.pandemic_felt_feelings == "A little worse" |
           gad.pandemic_felt_feelings == "A little better" |
           gad.pandemic_felt_feelings == "Much better")

whole_retro_prepan_scatter %>% ggplot(aes(x = gad.sum_score_retro, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4)

#Correlation little/much worse prepan retro gad

```{r Whole sample Correlation prep chunk}

# Check the items are the same from the whole chunk
prepan_retro_items

#Create a data set with only the GLAD and EDGI
prepan_retro_dat_worse <- dat[str_detect(dat$Sample, "GLAD|EDGI"), ]

#Filter the data to only include those with self-reported little/much worse gad since pandemic

prepan_retro_dat_worse <- prepan_retro_dat_worse %>% 
  filter(gad.pandemic_felt_feelings == "A little worse" |
           gad.pandemic_felt_feelings == "Much worse")

#Check that RAMP and NBR has been removed
prepan_retro_dat_worse %>% 
  freq(Sample)

# Calculate correlation matrix on specific columns (variables) from your data frame
prepan_retro.corr.mat_worse <- hetcor(
  as.data.frame(prepan_retro_dat_worse[,prepan_retro_items]), # Specify the data
  use = "pairwise.complete.obs" # Use all pairwise observations
)

# Print the correlation matrix
print(prepan_retro.corr.mat_worse, digits = max(3, getOption("digits") - 3))

# Save the correlations from the correlation matrix in an object
prepan_retro.corr.matrix_worse <- as.matrix(prepan_retro.corr.mat_worse)

# Save the p values from the correlation matrix in an object
prepan_retro.p.matrix_worse <- as.matrix(prepan_retro.corr.mat_worse$tests)

res1 <- cor.ci(prepan_retro.corr.matrix_worse)
lowci <- cor.plot.upperLowerCi(res1)
lowci[upper.tri(lowci)] = t(lowci)[upper.tri(lowci)]

upci <- cor.plot.upperLowerCi(res1)
upci[lower.tri(upci)] = t(upci)[lower.tri(lowci)]

conf <- paste0("[", format(lowci, digits=2), " - ", format(upci, digits=2), "]")

xs <- row(lowci)
ys <- (ncol(lowci)+1) - col(lowci)

xs[lower.tri(xs,diag=TRUE)] <- 0
ys[lower.tri(ys,diag=TRUE)] <- 0

```{r Little/much worse correlation matrix of prepan retro sum scores}
corrplot(prepan_retro.corr.matrix_worse, 
         method = "color", # objects to represent the correlations on plot
         type = "lower", # only use the lower triangle of the matrix
         diag = FALSE, # do not show the correlations on the diagonal
         addgrid.col = NA,
         addCoef.col = "black", # colour for the correlation coefficients in the plot
         tl.cex = 0.8,
         tl.col = "black",
         col=colorRampPalette(c("dodgerblue4","white","firebrick4"))(200), # colours for correlations
         # Combine with significance
         p.mat = prepan_retro.p.matrix, # Matrix with p values
         sig.level = 0.01, # Choose significant level
         insig = "blank" # Nonsignificant correlations have no colour
         )

text(xs, ys, conf, pos=1, cex=0.8)

```{r Little/much worse scatterplots for gad}

#Filter by little/much worse self-reported change since pandemic started in gad

much_worse_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "Much worse")

little_worse_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "A little worse")

much_worse_little_worse_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "Much worse" | 
           gad.pandemic_felt_feelings == "A little worse")

#Visualise each as a scatterplot

much_worse_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemic and retrospective GAD scores\nfor participants reporting much worse",
                      sep=" "),
        subtitle = paste("n", dim(much_worse_gad)[1]))

little_worse_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemic and retrospective GAD scores\nfor participants reporting a little worse",
                      sep=" "),
        subtitle = paste("n", dim(little_worse_gad)[1]))

much_worse_little_worse_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemic and retrospective GAD scores\nfor participants reporting much worse and a little worse",
                      sep=" "),
        subtitle = paste("n", dim(much_worse_little_worse_gad)[1]))

```{r Little/much better scatterplots for gad}

#Filter by participants reporting little/much better in gad since pandemic started (separetely and grouped)

little_better_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "A little better")

much_better_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "Much better")

much_better_little_better_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "Much better" | 
           gad.pandemic_felt_feelings == "A little better")

#Visualise them as scatterplots

little_better_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemic and retrospective GAD scores\nfor participants reporting a little better",
                      sep=" "),
        subtitle = paste("n", dim(little_better_gad)[1]))

much_better_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemic and retrospective GAD scores\nfor participants reporting a little worse",
                      sep=" "),
        subtitle = paste("n", dim(much_better_gad)[1]))

much_better_little_better_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") + 
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemic and retrospective GAD scores\nfor participants reporting much better and a little better",
                      sep=" "),
        subtitle = paste("n", dim(much_better_little_better_gad)[1]))
